Future cosmic microwave background delensing with galaxy surveys

### 2017-11-27 Notes/referee_response.tex

\vspace{0.4cm}

We agree that the phrasing of the sentences where secondary B-modes sources are introduced can be confusing. While it is well known in the field, it is  important to be clear to a broader audience.
For this reason we add this sentence in the introduction:

\textit{Two main effects produce non-primordial B-modes: the polarized foregrounds from our own galaxy  and the effect of gravitational interactions with large scale structures (LSS). 
This two components needs to be treated with very different techniques and in this work we will focus on the latter (see (CORE cleaning paper) for a review on the former).}

\vspace{1cm}
\item The author introduces sharp cuts to CIB and WISE at $\ell < 100$. Is the
alternative approach of incorporating the foreground content in the
auto spectra of the tracer (E.g. Eq. 17) combination possible ? That
would retain some more of the signal presumably. 
\vspace{0.4cm}
This is a good idea. The real obstacle to that is our current knowledge of foreground spectra. There is not a trustworthy model for that. However one might measure that empirically by measuring $C^{II}$ and $C^{I\phi}$ directly from the data. Doing that will call everything in the tracer map that does not correlate with $\phi$, "noise". That noise will include foregrounds too. However..
Given the small amount of delensing power for $\ell<100$ the field decided, so far, to cut those in forecasts. Given how much foregrounds are complicated I believed only data will answer this. An exploration of these on simulations is probably beyond the scope of this paper, but very soon future delensing work will need to explore this on real data.

### 2017-11-28 Notes/referee_response.tex

Furthermore, while I agree that the improvement would be even more significant for T-E delensing, it is nonetheless also quite important for BB too.
As can be seen in Figure 4 in the paper LSST shows a significant improvement in the correlation with the lensing potential all the way to $\ell\sim 800$. Even for other probes, in general, less powerful like, there is still an improvement for example where the contribution to BB (the dashed line peaks at $\ell\sim350$).
Below I am attaching a table that shows the improvement in the BB power removed and constraints on $\sigma(r)$.

\begin{table}[htp]
\caption{Improvement on $\sigma(r)$ constraints together with the BB power removed (in parenthesis) with and without binning.}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Survey & No Binning & Binning\\
\hline
DES & 1.17 (14\%)& 1.20 (16\%) \\
\hline
LSST & 1.67 (40\%) & 2.10 (52\%)\\
\hline
DESI & 1.10 (9\%) & 1.13 (12\%)  \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

\vspace{1cm}

\item
The author uses iterative internal delensing noise levels for the
CMB. I believe this could be discussed in more details. Can he specify
what multipole range is considered for this calculation ? I suppose
this multipole range should not overlap with the primordial B that is
used for constraints on r, however, this range (unnecessarily) extends
up to 800 in the S3 section for example. Most likely, 200 would be
largely enough, does it imply a sizable reduction on the iterative
reconstruction noise levels ? Besides, iterative delensing is
complicated and not established yet, in contrast to Quadratic
Estimator delensing. The paper might gain in interest if numbers for
Q.E. and iterative delensing were presented.

I

### 2017-11-28 Notes/referee_response.tex

\title{  \scriptsize{Response to Referee for: "Future cosmic microwave background delensing with galaxy surveys"}}
\author{A. Manzotti}
\date{} % delete this line to display the current date

%%% BEGIN DOCUMENT
\begin{document}

\maketitle

I want to thank the referee for the very careful review of the paper, and for the comments, corrections and suggestions that ensued.

@@

\vspace{1cm}

\item
The author uses iterative internal delensing noise levels for the
CMB. I believe this could be discussed in more details. Can he specify
what multipole range is considered for this calculation ? I suppose
this multipole range should not overlap with the primordial B that is
used for constraints on r, however, this range (unnecessarily) extends
up to 800 in the S3 section for example. Most likely, 200 would be
largely enough, does it imply a sizable reduction on the iterative
reconstruction noise levels ? Besides, iterative delensing is
complicated and not established yet, in contrast to Quadratic
Estimator delensing. The paper might gain in interest if numbers for
Q.E. and iterative delensing were presented.

\vspace{0.4cm}

One information that is probably not clear enough or partially missing in the text is that iterative delensing compared to iterative delensing is important only for CMB-S4 noise levels. For example, the amount of BB power removed for CMB-S3 level using iterative vs. QE is 56\% vs. 58\%.

Furthermore, the idea of a separation in scale between the CMB data used to reconstruct the potential and the data used to constrained r, to solve the so-called "internal delensing bias" is important for both QE and iterative approach. 
While important, we left a full study...

In some sense, we assume that either the technique applied in Namikava 2017 or () can work without a significant loss of signal. (put that in the text?s)
Again to test this one needs either real data or full ray tracing simulation for CMB lensing. For example, the deep 100 deg$^{2}$ field measured by SPTPol that can reach a depth of 5-6 muK arcmin with high resolution might be the first dataset where this can be explored.

Following the referee comments, we
\begin{itemize}
\item Make it clear that only CMB S4 will benefit from iterative delensing.
\item Add the level of delensing that is possible to achieve with CMB-S4 if the well tested QE is used instead of the iterative approach. This gives in some sense a range of possible efficiency for CMB S4 experiments.
\end{itemize}

### 2017-11-28 Notes/galaxy_delensing.tex
working on referee response

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}.

@@

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.

>>>>

The CMB lensing potential is \textit{the} field that we need in order to reverse the effect of large scale structure and delens the CMB.
However, the lensing potential can also be reconstructed using the CMB itself. In that case, we can treat it as a noisy tracer of the true field. Both the CMB lensing field and its noisy reconstructed counterpart have the same kernel $W^{\kappa}(z)$.
However, when computing the power spectrum of the latter, we need to add a noise component. Given the instrumental noise level and the beam, we can calculate the reconstruction noise $N_{\ell}^{\kappa \kappa}$, and so
\beq\label{eqn:noisekappa}
\cl^{\kappa_\mathrm{rec} \kappa_\mathrm{rec}} = \cl^{\kappa \kappa } +N^{\kappa \kappa}_{\ell}
\eeq
In this work, the level of noise is computed assuming an iterative approach to the CMB lensing reconstruction as described in \cite{smith:2012,hirata:2003}. The iterative approach has not been fully demonstrated on data yet but it is among the major goals of the CMB community (\cite{abazajian:2016}).
Note however that, as pointed out in in \cite{smith:2012,hirata:2003}, this improves the CMB lensing reconstruction compared to a quadratic estimator approach only for CMB-S4 noise levels.

@@

The correlation factor attainable using generation 3 experiments is shown in \reffig{corrS3}.
An interesting finding is that DESI will be less efficient (removing $11\%$ of the power) than a DES-like survey ($17\%$) despite the fact that it can probe slightly higher redshift. The reason is that, because of the broad CMB kernel, spectroscopic redshift accuracy is not needed and the lower shot noise in DES increases the delensing efficiency. Unfortunately, adding DESI will only bring the power removed using LSS from 41 to 45\%.
Contrary to current S2 experiments, the CMB internal reconstruction will dominate the correlation with the lensing potential up to $\ell\simeq 600$.
For CMB S3, the improvement due to an iterative approach compared to the quadratic estimator is still minimal so that the more traditional technique can be used. For example, in this case, the removed power goes from $56\%$ to $54\%$ (quoted in \reftab{S3}) when using a quadratic estimator instead of the iterative one.
